Narrative

### 2023-05-21 manuscript/methods.qmd
method: improved parameter section

We ran our analyses across a range of within-study sample sizes $N$ and number of studies $K$. Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$. We made this choice rather than treating sample size as a random variable (as done by e.g., REF1, REF2) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance. We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology by XXX (still needs lookup) and consider the following values {VALUES}.

When focusing on the bias of heterogeneity estimate, the value of $K$ would not matter for an unbiased estimator. However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (REF Viechtbauer). Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance. As such, the values of $K$ can affect bias, especially at lower levels of heterogeneity. In their review of meta-analyses in psychology, van Erp et al. (YEAR) find $K$ values ranging from XX to XX, and in our simulations we consider the following values {VALUES}.

We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the interaction between effect size and reliability variance. For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale. For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al., (REF) and Schäfer & Scharz (REF). Nuijten et al. performed a meta-meta-analysis on the fields of RESEARCH and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies. This estimate is likely positively biased due to publication bias and selective reporting. Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in FIELD, with an 'upper median' (i.e., the 83.35% quantile) of 0.41. This estimate is likely less affected by positive bias but should not be expected to be unbiased (REF Scheel).

We defined the between-studies standard deviation ($\tau$) to cover a wide range of variability in true effect sizes and empirical estimates reported in the psychological literature. Van Erp et al., (REF) provide empirical heterogeneity estimates from XX meta-analyses in XX article published in the journal Psychological Bulletine between 1990 - 2013. The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24). The point of this study is that empirical heterogeneity estimates may not be accurately estimated. Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al. for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them reasonable variability on the product-moment correlation scale. The values we selected were {0, 0.1, 0.15, 0.2}. We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1. Our values are similar to those used by Brannick et al. (2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau = $ 0, 0.08, 0.13, 0.2).

Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale. To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale. The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (REF Higgins), and as such is measured from 0 - 100%. The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$. This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (REF: Higgins, I2 is not an absolute measure of heterogeneity). However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.

We thus computed the $I^2$ index for each sample size condition and $\tau$-values defined in Pearson's $r$, and then computed corresponding $\tau$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes. Two complications were that the value for $\sigma^2$ varies with effect size for Pearson's $r$ and that effect size varies across studies because of $\tau$. We resolved this by setting $\theta = 0$ and computing the expected $\sigma^2$ given $\tau$ and sample size and then using this expected value to compute $I^2$. We used the 'law of the unconscious statistician' to compute the expected value of $\sigma^2$, which says that the expected value of a function $g(X)$ of a random variable can be expressed in terms of the probability distribution of X: $E[g(X)] = \int_{-\infty}^\infty g(x) f(x) dx$. In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999. The resulting $\tau$ values for Fisher $z$ are only minimally different from {0, 0.1, 0.15, 0.2}. For example, for $N = 150$ the corresponding Fisher $z$ $\tau$-values were {VALUES} and as $N$ increased heterogeneity values were more similar (see supplement XX).

Finally, we defined average reliability levels and their standard deviation based on empirical estimates from the literature. Flake et al. (YEAR) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13. The interquartile range was approximately 0.68 - 0.87 for studies wherein the associated scale lacked a reference and 0.79 - 0.88 for scales that did have a reference. Sanchez-Meca et al. (REF) report the reliability estimates based on five 'reliability generalization studies'. These five studies summarized reliability in 25 - 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 - 0.133. Given some likely positive bias in such empirical values, we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.

>>>>

We ran our analyses across a range of within-study sample sizes $N$ and number of studies $K$. Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$. We made this choice rather than treating sample size as a random variable (as done by e.g., Field, 2005; Brannick et al., 2019) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance. We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology. Fraley et al., report median sample sizes for between-person studies in 9 psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year. Sassenber and Ditrich (2019) find a median sample size of 110 (interquartile range, 71 - 195) across 4 journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al. (personal communication, 2023) find median sample sizes between 24 - 184 for 6 journals in psychology for the years 1995, 2006, 2019. Given these empirical findings we consider the following sample size values {50, 100, 150, 200}.

When focusing on the bias of heterogeneity estimate, the number of meta-analyzed studies, $K$, would not matter for an unbiased estimator. However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (Viechtbauer, 2005). Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance in $\tau^2$. As such, values of $K$ can affect bias, especially at lower levels of heterogeneity. We consider the following values of $K$: {5, 20, 40, 100}.

We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the interaction between effect size and reliability variance. For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale. For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al. (2020) and Schäfer & Scharz (2019). Nuijten et al. performed a meta-meta-analysis on the fields of intelligence research and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies. This estimate is likely positively biased due to publication bias and selective reporting. Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in psychology in general, with an 'upper median' (i.e., the 83.35% quantile) of 0.41. This estimate is likely less affected by positive bias but should not be expected to be unbiased as there appears to be some positive bias also in preregistered research on average (Scheel, 2021).

We defined the between-studies standard deviation ($\tau$) to cover a wide range of variability in true effect sizes and empirical estimates reported in the psychological literature. Van Erp et al. (2017) provide empirical heterogeneity estimates from 747 meta-analyses in 61 article published in the journal Psychological Bulletin between 1990 - 2013. The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24). The point of this study is that empirical heterogeneity estimates may not be accurately estimated. Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al. for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them to represent reasonable variability on the product-moment correlation scale. The values we selected were {0, 0.1, 0.15, 0.2}. We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1. Our values are similar to those used by Brannick et al. (2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau = $ 0, 0.08, 0.13, 0.2).

Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale. To compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale. The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (Higgins, 2002; 2003), and as such is measured from 0 - 100%. The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$. This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (Borenstein, 2017). However, because we keep $N_i$ fixed across studies, the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.

We thus computed the $I^2$ index for each sample size condition and $\tau$-values defined in Pearson's $r$, and then computed corresponding $\tau$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes. Two complications were that the value for $\sigma^2$ varies with effect size for Pearson's $r$ and that effect size varies across studies because of $\tau$. We resolved this by setting $\theta = 0$ and computing the expected $\sigma^2$ given $\tau$ and sample size and then using this expected value to compute $I^2$. We used the 'law of the unconscious statistician' to compute the expected value of $\sigma^2$, which says that the expected value of a function $g(X)$ of a random variable can be expressed in terms of the probability distribution of X: $E[g(X)] = \int_{-\infty}^\infty g(x) f(x) dx$. In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999. The resulting $\tau$ values for Fisher $z$ were only minimally different from {0, 0.1, 0.15, 0.2}. For example, for $N = 50$ the corresponding Fisher $z$ $\tau$-values were {0, 0.1031263, 0.1566007, 0.2123514} and as $N$ increased heterogeneity values were more similar (see supplement XX).

Finally, we defined average reliability levels and their standard deviation based on empirical estimates from the literature. Flake et al. (2017) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13. The interquartile range was approximately 0.68 - 0.87 for studies wherein the associated scale lacked a reference and 0.79 - 0.88 for scales that did have a reference. Sánchez-Meca et al. (2003) report the reliability estimates based on five 'reliability generalization studies'. These five studies summarized reliability in 25 - 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 - 0.133. Given some likely positive bias in such empirical values (Hussey et al., 2023), we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.
